Plots_manuscript_v0.3

### 2020-07-30 R/Plots_manuscript_v0.3.Rmd
Add BLUPs for subpop specific kinship GWAS

h2all %>%
  filter(site %in% c("eight_sites")) %>%
  mutate(GandGxE = Estimate + GxE_Estimate)  %>%
  select(GandGxE, site:phe, Estimate, GxE_Estimate)
```

@@

# D2F

@@

## G_BLUP GWAS

>>>>

h2all %>%
  filter(site %in% c("eight_sites")) %>%
  mutate(GandGxE = Estimate + GxE_Estimate)  %>%
  select(GandGxE, site:phe, Estimate, SE, GxE_Estimate, GxE_SE)
```

@@

# ----
# D2F

@@

# ----------------------------
# G_BLUP GWAS

The plan is to do GWAS on G_BLUP from an eight site model with GxE modeled as site:kinship for each subpopulation and both combined. The traits I will do GWAS on for G_BLUPs are as follows:

eight_sites
FL50 (to show other assoc are stronger)
dyln_fl50
cgdd_12c_gr2fl

Maybe also do GWAS on G_BLUP from the North and Texas models as above, for the traits:
tx_sites & north_sites
FL50 (to show other assoc are stronger)
dyln_fl50
cgdd_12c_gr2fl
crain_gr2fl
crain_1d
GR50
cgdd_12c_10d

All I really want to show are that there are stronger G associations for the weather-derived variables than for FL50 and GR50. And possibly exploring G associations at subpopulation home and away sites, for the second bit. 
That might be too complicated. Eh.

It's possible I might want to confirm some of these G candidates with the fourway - how we see similar patterns in this cross for FL50 vs dyln/GDD, and there are some similar associations.

THEN, for the next section, do GWAS on the BLUPs from the eight single site models. For these traits:
all single sites...
FL50 (actually, maybe not)
dyln_fl50
cgdd_12c_gr2fl
crain_gr2fl
crain_1d
--
GR50 (or maybe not)
cgdd_12c_10d

Then do mash runs. First for single traits across all eight sites, to make U_hyp for the next step - take the U_ed with large percentage variance explained. Make U_hyp with those covariances but no covariance with anything else... 
Then run all the flowering-related traits together. Talk about how then SNPs can have effects that are strongest in response to one type of environmental cue. And how cues can be present or absent at different sites. And how this kind of analysis can help reveal all the patterns at play, because it doesn't make assumptions about how effects are structured.

Maybe bring back the QTL maps after this and talk about how patterns of effects differ across sites in these models, too. Though the Genstat model is less sensitive...

```{r}
h2all %>%
  filter(site %in% c("eight_sites"))
```

## Make phenotype df for GWAS
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_eight <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl")
phe_txano <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_lists <- list(TX1 = phe_single, TX2 = phe_single, TX3 = phe_single, OK = phe_single, MO = phe_single,
               NE = phe_single, MI = phe_single, SD = phe_single, eight_sites = phe_eight, tx_sites = phe_txano, north_sites = phe_txano)

i = 9
j = 1
k = 1
phe_gwas <- tibble()

for(j in seq_along(subpop_v)){
    for(i in seq_along(site_v)){
      phe_v <- phe_lists[[i]]
      for(k in seq_along(phe_v)){
        phe_raw <- read_csv(file.path(outputdir, names(site_v)[i],
                                      paste0("blups_", names(site_v)[i], "_",
                                             names(subpop_v)[j], "_", phe_v[k],
                                             ".csv")))
        names(phe_raw)[2] <- paste0(phe_v[k], "_", names(site_v)[i])
        phe_proc <- phe_raw %>%
          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
          filter(!is.na(PLANT_ID) & !grepl("^SITE", factor)) %>%
          left_join(metadata, by = "PLANT_ID") %>%
          filter(SUBPOP %in% subpop_v[[j]]) %>%
          select(PLANT_ID, 3)
        if(i == 1 & k == 1){
          phe_gwas <- phe_proc
        } else {
          phe_gwas <- phe_gwas %>%
            left_join(phe_proc, by = "PLANT_ID")
        }
      }
    }
  phe_gwas <- phe_gwas %>%
  select(PLANT_ID, ends_with("eight_sites"), ends_with("tx_sites"), ends_with("north_sites"), everything())
  write_csv(phe_gwas, path = file.path(workingdir, "analysis", "gwas", "BLUPs",
                                       paste0("BLUP_phenotypes_",
                                              "for_gwas_GR_FL_",
                                              names(subpop_v)[j],
                                              ".csv")))
}

ggplot(data = phe_gwas, aes(x = phe_gwas$cgdd_12c_gr2fl_eight_sites)) + geom_histogram()

## Run GWAS 
```{r}
library(bigsnpr)
library(tidyverse)
library(switchgrassGWAS)
library(AnnotationDbi)
txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
inputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs")
subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")
snp_v <- c("12\\.1M", "8\\.8M", "10\\.3M")

inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")
set_v <- list(eight_sites = "eight_sites", tx_sites = "tx_sites", north_sites = "north_sites", single_sites = "!sites")

for(i in seq_along(subpop_v)){
  for(j in seq_along(set_v)){
    dir.create(file.path(outputdir, names(set_v)[j]), showWarnings = FALSE)
    snp <- snp_attach(inputfiles$SNPfiles[i])
    phe_df <- read_csv(file.path(inputdir, paste0("BLUP_phenotypes_",
                                                  "for_gwas_GR_FL_",
                                                  names(subpop_v)[i],
                                                  ".csv"))) 
  if(names(set_v)[j] != "single_sites"){
    phe_df <- phe_df %>%
    select(PLANT_ID, ends_with(names(set_v)[j]))
  } else {
    phe_df <- phe_df %>%
    select(PLANT_ID, !ends_with("sites"))
  }
  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", 
                      outputdir = file.path(outputdir, names(set_v)[j]),
                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
                      minphe = 130)

  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir,
                                                       names(set_v)[j]), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       snp_v[i]))
  phenotypes <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
  out1 <- pvdiv_bigsnp2mashr(path = file.path(outputdir, names(set_v)[j]), 
                             snp = snp, gwas_rds = gwas_rds, 
                             phenotypes = phenotypes, numSNPs = 1000, 
                             clump = FALSE, scaled = FALSE, saveoutput = TRUE,
                             suffix = paste0("FL_", names(set_v)[j], "_",
                                             names(subpop_v)[i]))
  }
}
